\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-02-01 paper.qmd
Changed the table for a reproducable one

library(grid)
# library(xtable)
# 
# xtable2kable <- function(x) {
#   out <- capture.output(print(x, table.placement = NULL))[-(1:2)]
#   out <- paste(out, collapse = "\n")
#   structure(out, format = "latex", class = "knitr_kable")
# }

@@

# error_df <- read_csv("data/pbmc/pbmc_3k_festem/error_pbmc_umap.csv")
# 
# error_df |>
#   knitr::kable() 
```

| n_neighbors | min_dist | metric     | error                 |
|-------------|----------|------------|-----------------------|
| **30**      | **0.3**  | **cosine** | **38634.66719286174** |
| 5           | 0.99     | euclidean  | 35938.98594054407     |
| 5           | 0.99     | cosine     | 35597.01981261912     |
| 15          | 0.99     | euclidean  | 35534.75778935427     |
| 30          | 0.99     | euclidean  | 35664.68763936665     |
| 1022        | 0.99     | cosine     | 37022.0125249776      |
| 1022        | 0.3      | cosine     | 39746.347504639125    |
| 800         | 0.3      | euclidean  | 38385.25485420897     |
| 100         | 0.99     | euclidean  | 36528.57424747017     |
| 230         | 0.99     | cosine     | 37864.26802578186     |
| 354         | 0.99     | euclidean  | 37530.86598361975     |

@@

plot_list2_pbmc_error <- error_df |>
  mutate(type = factor(type , levels = c("error 2 or less", "error 2-10", "error 10-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error greter than 35"))) |>
    ggplot(aes(x = UMAP1,
               y = UMAP2, color = type))+
    geom_point(alpha=0.5, size = 0.1) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#fee0d2", "#fcbba1",
                                         "#fc9272", "#fb6a4a", "#ef3b2c",
                                         "#cb181d", "#a50f15", "#99000d")) #+
  #geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
  #annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
```

@@

<!--tSNE parm 1-->

tSNE_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_tsne_5.rds")

```{r}
#| warning: false
#| echo: false
num_bins_pbmc <- calculate_effective_x_bins(.data = tSNE_pbmc, x = tSNE1,
                                            cell_area = 1) ##23

shape_value_pbmc <- calculate_effective_shape_value(.data = tSNE_pbmc,
                                                    x = tSNE1, y = tSNE2) ## 0.8772751
shape_value_pbmc <- 0.8772751

## To extract bin centroids
hexbin_data_object_pbmc <- extract_hexbin_centroids(nldr_df = tSNE_pbmc,
                                                    num_bins = num_bins_pbmc,
                                                    shape_val = shape_value_pbmc, x = tSNE1, y = tSNE2)

tSNE_pbmc_with_hb_id <- tSNE_pbmc |>
  dplyr::mutate(hb_id = hexbin_data_object_pbmc$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), tSNE_pbmc_with_hb_id)

## To find the benchmark value
benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)

trimesh_removed_pbmc_tsne <- remove_long_edges(.data = distance_pbmc, benchmark_value = benchmark_pbmc,
                                               triangular_object = tr1_object_pbmc, distance_col = distance)

trimesh_removed_pbmc_tsne <- trimesh_removed_pbmc_tsne +
  geom_point(size = 2, colour = "#33a02c") +
  #ggtitle("(b)") +
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
  #coord_equal() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  )

plot_list2_pbmc_error <- error_df |>
    mutate(type = factor(type , levels = c("error 2 or less", "error 2-10", "error 10-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error greter than 35"))) |>
    ggplot(aes(x = tSNE1,
               y = tSNE2, color = type))+
    geom_point(alpha=0.5, size = 0.1) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#fee0d2", "#fcbba1",
                                         "#fc9272", "#fb6a4a", "#ef3b2c",
                                         "#cb181d", "#a50f15", "#99000d"))

```{r}
#| echo: false
#| fig-cap: "(a) 2D layout from tSNE (perplexity = 5) applied for the PBMC3k dataset. Is this a best representation of the original data?, (b)  Model in the 2D space with tSNE (<>)"
#| label: fig-nldervisPBMCtSNE1
#| out-width: 100%
#| fig-pos: h

#plot_list2_pbmc + 
trimesh_removed_pbmc_tsne + plot_list2_pbmc_error +
plot_layout(guides='collect', ncol=3) &
  theme(legend.position='bottom')

<!--tSNE parm 2-->

tSNE_pbmc <- read_rds("data/pbmc/pbmc_3k_festem/pbmc_tsne_2.rds")

```{r}
#| warning: false
#| echo: false
num_bins_pbmc <- calculate_effective_x_bins(.data = tSNE_pbmc, x = tSNE1,
                                            cell_area = 1) ##23

shape_value_pbmc <- calculate_effective_shape_value(.data = tSNE_pbmc,
                                                    x = tSNE1, y = tSNE2) ## 0.8772751
shape_value_pbmc <- 0.8772751

## To extract bin centroids
hexbin_data_object_pbmc <- extract_hexbin_centroids(nldr_df = tSNE_pbmc,
                                                    num_bins = num_bins_pbmc,
                                                    shape_val = shape_value_pbmc, x = tSNE1, y = tSNE2)

tSNE_pbmc_with_hb_id <- tSNE_pbmc |>
  dplyr::mutate(hb_id = hexbin_data_object_pbmc$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_pbmc <- dplyr::bind_cols(training_data_pbmc |> dplyr::select(-ID), tSNE_pbmc_with_hb_id)

## To find the benchmark value
benchmark_pbmc <- find_benchmark_value(.data = distance_pbmc, distance_col = distance)

trimesh_removed_pbmc_tsne <- remove_long_edges(.data = distance_pbmc, benchmark_value = benchmark_pbmc,
                                               triangular_object = tr1_object_pbmc, distance_col = distance)

trimesh_removed_pbmc_tsne <- trimesh_removed_pbmc_tsne +
  geom_point(size = 2, colour = "#33a02c") +
  #ggtitle("(b)") +
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
  #coord_equal() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  )

plot_list2_pbmc_error <- error_df |>
    mutate(type = factor(type , levels = c("error 2 or less", "error 2-10", "error 10-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error greter than 35"))) |>
    ggplot(aes(x = tSNE1,
               y = tSNE2, color = type))+
    geom_point(alpha=0.5, size = 0.1) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#fee0d2", "#fcbba1",
                                         "#fc9272", "#fb6a4a", "#ef3b2c",
                                         "#cb181d", "#a50f15", "#99000d"))

```{r}
#| echo: false
#| fig-cap: "(a) 2D layout from tSNE (perplexity = 2) applied for the PBMC3k dataset. Is this a best representation of the original data?, (b)  Model in the 2D space with tSNE (<>)"
#| label: fig-nldervisPBMCtSNE2
#| out-width: 100%
#| fig-pos: h

#plot_list2_pbmc + 
trimesh_removed_pbmc_tsne + plot_list2_pbmc_error +
plot_layout(guides='collect', ncol=3) &
  theme(legend.position='bottom')

>>>>

library(grid)
library(xtable)

xtable2kable <- function(x) {
  out <- capture.output(print(x, table.placement = NULL))[-(1:2)]
  out <- paste(out, collapse = "\n")
  structure(out, format = "latex", class = "knitr_kable")
}

@@

<!--https://dicook.github.io/mulgar_book/7-spin-and-brush.html#fig-penguins-bs-detourr-->

error_df <- read_csv("data/pbmc/pbmc_3k_festem/error_pbmc_umap.csv", col_names = FALSE)

error_df <- error_df |>
  mutate(X4 = round(X4/100, 2)) |>
  arrange(X4)

colnames(error_df) <- c("n_neighbors", "min_dist", "metric", "error (x 100)")

xtable(error_df, caption = "errors for different parameter combinations", label = "tbl-errordiffcomb") |>
  xtable2kable() |>
  kable_styling(latex_options = c("striped", "HOLD_position"), position = "center", full_width = FALSE) |>
  row_spec(0, bold = TRUE) |>
  row_spec(3, bold=T, color = "red")
```

@@

plot_list2_pbmc_error <- error_df |>
  mutate(type = factor(type , levels = c("error 2 or less", "error 2-10", "error 10-15", "error 15-20", "error 20-25", "error 25-30", "error 30-35", "error greter than 35"))) |>
    ggplot(aes(x = UMAP1,
               y = UMAP2, color = type))+
    geom_point(alpha=0.5, size = 0.1) +
    coord_equal() +
    theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold")) + #ggtitle("(a)") +
  theme_linedraw() +
    theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
        legend.title = element_text(size=5), #change legend title font size
        legend.text = element_text(size=4),
         legend.key.height = unit(0.25, 'cm'),
         legend.key.width = unit(0.25, 'cm')) +
  scale_color_manual(values=c("#fee0d2", "#fcbba1",
                                         "#fc9272", "#fb6a4a", "#ef3b2c",
                                         "#cb181d", "#a50f15", "#99000d")) #+
  #geom_text(aes(x=UMAP1,y = UMAP2,label = cell_label), data = class_avg,inherit.aes = F, color = "black",fontface = "bold",size = 2) +
  #annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

plot_list2_pbmc + trimesh_removed_pbmc_umap + plot_list2_pbmc_error +
plot_layout(guides='collect', ncol=3) &
  theme(legend.position='bottom')

### 2024-02-01 paper.tex
Changed the table for a reproducable one

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, enhanced, sharp corners, breakable, interior hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

\begin{longtable}[]{@{}llll@{}}
\toprule\noalign{}
n\_neighbors & min\_dist & metric & error \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{30} & \textbf{0.3} & \textbf{cosine} &
\textbf{38634.66719286174} \\
5 & 0.99 & euclidean & 35938.98594054407 \\
5 & 0.99 & cosine & 35597.01981261912 \\
15 & 0.99 & euclidean & 35534.75778935427 \\
30 & 0.99 & euclidean & 35664.68763936665 \\
1022 & 0.99 & cosine & 37022.0125249776 \\
1022 & 0.3 & cosine & 39746.347504639125 \\
800 & 0.3 & euclidean & 38385.25485420897 \\
100 & 0.99 & euclidean & 36528.57424747017 \\
230 & 0.99 & cosine & 37864.26802578186 \\
354 & 0.99 & euclidean & 37530.86598361975 \\
\end{longtable}

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisPBMCtSNE1-1.pdf}

\caption{\label{fig-nldervisPBMCtSNE1}(a) 2D layout from tSNE
(perplexity = 5) applied for the PBMC3k dataset. Is this a best
representation of the original data?, (b) Model in the 2D space with
tSNE (\textless\textgreater)}

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-nldervisPBMCtSNE2-1.pdf}

\caption{\label{fig-nldervisPBMCtSNE2}(a) 2D layout from tSNE
(perplexity = 2) applied for the PBMC3k dataset. Is this a best
representation of the original data?, (b) Model in the 2D space with
tSNE (\textless\textgreater)}

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, frame hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi

@@

\begin{table}[H]
\centering
\begin{tabular}{rrrlr}
  \hline
\textbf{} & \textbf{n\_neighbors} & \textbf{min\_dist} & \textbf{metric} & \textbf{error (x 100) }\\ 
  \hline
\cellcolor{gray!6}{1} & \cellcolor{gray!6}{5.00} & \cellcolor{gray!6}{0.99} & \cellcolor{gray!6}{cosine} & \cellcolor{gray!6}{355.97 }\\ 
  2 & 230.00 & 0.99 & cosine & 378.64 \\ 
\textcolor{red}{\textbf{\cellcolor{gray!6}{  3}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{30.00}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{0.30}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{cosine}}} & \textcolor{red}{\textbf{\cellcolor{gray!6}{386.35 }}}\\ 
   \hline
\end{tabular}
\caption{errors for different parameter combinations} 
\label{tbl-errordiffcomb}
\end{table}

@@
